---
title: "CHAPTER 8: SIMULATIONS IN R"
output: html_notebook
---

8.1 Mathematics Function
In Chapter 1, it was shown that R can function like a calculator. It is also able to be used to solve complex mathematical problems
including calculus and linear algebra. Some of built-in mathematics functions are as follows.


Example 8.1a Minimun and Maximum
```{r}
## Min, Max, pmin, pmax
v1 <- c(5,3,11,7,6,8)
m1 <- matrix(v1, nrow=3, ncol=2)
m1
```

```{r}
min(m1[,1],m1[,2])
min(m1[,1])
min(m1[,2])
```

```{r}
pmin(m1[,1],m1[,2])
```

```{r}
max(m1[,1],m1[,2])
```

```{r}
pmax(m1[,1],m1[,2]) 
```

```{r}
colMeans(m1)
```

```{r}
rowMeans(m1)
```

```{r}
sqrt(var(m1[,1]))
sqrt(var(m1[,2]))
```



Example 8.1b Cumulative Sum and Product
```{r}
## Cumulative Sum and Product
x <- seq(2, 8, by = 2)
x
```

```{r}
sum(x)
```

```{r}
cumsum(x)
```

```{r}
prod(x)
```

```{r}
cumprod(x)
```

```{r}
#own codes
zm <- seq(-9, 18, by = 3)
zm
sum(zm)
cumsum(zm)
prod(zm)
cumprod(zm)
```


Example 8.1c Minimization/ Maximization Function
```{r}
# Function to Minimize/ Maximize: nlm(.)
# Apply Newton-Raphson method
# nlm(function(input) return(equation), initial input)
x <- 1:1000
head(x);tail(x)
nlm(function(x) return(x^2 -2* sin(x)),10)

y <- x^2 - 2*sin(x)
head(y);tail(y)
plot(y)
```


```{r}
x2 <- seq(-10,10,by=0.2)
head(x2); tail(x2)
nlm(function(x2) return(x2^2 -2* sin(x2)),10)

y <- x2-2 +2* sin(x2)
head(y); tail(y)
plot(y)
```


Example 8.1d Calculus

Differentiation:
  D(expression(function)),"variable")
  
Definite Integration:
  integrate(function(argument) equation, lower limit, upper limit) 
```{r}
## Calculus
## Differentiation
D(expression(x + 2*x^3),"x")
```

```{r}
## Integration
integrate( function(x) sqrt(x) , 0,1)
```


8.2 Function for Statistical Distribution
R provides function for most of the statistical distributions. The
probability distribution functions are prefixed as follows.
• d : density or probability density function (pdf)
• p : cumulative density function (cdf)
• q : quantile function or inverse of cdf
• r : random number generation

Statistical Distribution Function in R

Normal Distribution

Example 8.2a Probability Density Function: dnorm()
```{r}
## dnorm, Syntax: dnorm(x, mean, sd)
x <- seq(-10, 10, by = .01)
y <- dnorm(x, mean = 5.5, sd = 0.5)
head(x);tail(x)
plot(x,y, main = "Normal Distribution", col = "pink")

```

```{r}
x <- seq(-10, 10, by = .01)
y <- dnorm(x, mean = 0, sd = 1)
plot(x,y, main = "Normal Distribution", col = "red")
```



Example 8.2b Cumulative Density Function: pnorm()
```{r}
#pnorm()
#Syntax: pnorm(x,mean,sd)
x <- seq(-10, 10, by = .01)
y <- pnorm(x, mean = 5.5, sd = 0.5)
plot(x,y, main = "pnorm()", col = "blue")
```


Example 8.2c Quantile Function: qnorm()
```{r}
#qnorm()
#Syntax: qnorm(x,mean,sd)
x <- seq(0, 1, by = 0.02)
y <- qnorm(x, mean = 2, sd = 1)
plot(x,y, main = "qnorm()", col = "pink")


###Note that, input value for this function must be between 0 and 1.
```


Example 8.2d Randon data: rnorm()

```{r}
set.seed(7035)
x_nonstandard <- rnorm(50,12,3) #n=50, mean=12, sd=3
x_nonstandard
describe(x_nonstandard)
plot(x_nonstandard, main="Normal dist. mean=12, std=3")
hist(x_nonstandard, main="Normal dist. mean=12, std=3")
```

```{r}
#rnorm()
#Syntax: rnorm(n, mean, sd)
y <- rnorm(50) #standard normal
hist(y, main = "Normal Distribution (mean=0, std.=1)",
col = "cyan")
```


Poisson Distribution

Example 8.2e Probability Density Function: dpois()
```{r}
## dpois, Syntax: dpois(x, lambda)
set.seed(57)
x <- seq(0, 50, by = 1); head(x); tail(x)
y <- dpois(x, lambda=3); head(x); tail(x)
plot(x,y, main = "Poisson PDF,lambda=3", col = "purple")

```


Example 8.2f Cumulative Density Function: ppois()
```{r}
#ppois()
#Syntax: ppois(x,lambda)
set.seed(57)
x <- seq(0, 50, by = 1)
y <- ppois(x, lambda=3)
plot(x,y, main = "Poisson CDF: ppois()", col = "lightpink3")

```


Example 8.2g Quantile Function: qpois()
```{r}
#qpois()
#Syntax: qpois(x,lambda)
set.seed(57)
x <- seq(0, 1, by = 0.01)
y <- qpois(x, lambda=3)
plot(x,y, main = "Poisson Quantile: qpois()", col = "maroon")

###Note that, input value for this function must be between 0 and 1.
```


Example 8.2h Random data: rpois()
```{r}
# random number: rpois(x,lambda)
set.seed(57)
dt_pois <- rpois(50,5)
head(dt_pois); tail(dt_pois)

library(psych)
describe(dt_pois)
par(mfrow=c(1,2))
plot(dt_pois, main = "Poisson Random: rpois()", col= "plum")
hist(dt_pois, main = "Histogram: Poisson Random: rpois()", col= "violet")
```


8.3 Simulation Programming in R

Example 8.3a Generate Random Numbers

For normal distribution, the random numbers can be generated as
follows:

• Generate random number of Standard normal
```{r}
## Generate random number of standard normal
x_standard <- rnorm(50) #n=50, mean=>0, sd=<1
x_standard

## Descriptive summary
library(psych)
describe(x_standard)
hist(x_standard, main="Normal dist. mean=0, std=1")

```

• Generate random number of non-standard normal
```{r}
x_nonstandard <- rnorm(50,12,3) #n=50, mean=12, sd=3
x_nonstandard
describe(x_nonstandard)
hist(x_nonstandard, main="Normal dist. mean=12, std=3")
```


Descriptive Summary and Histogram - Standard Normal

Descriptive Summary and Histogram - Non-Standard
Normal



Seed of Random Numbers

Example 8.3b set.seed()
```{r}
## seed of random number
set.seed(1234) #to reproduce the same group of values
x_standard <- rnorm(50) ; x_standard
describe(x_standard)
hist(x_standard, main="Normal dist. mean=0, std=1")
```

Example 8.3c set.seed()
```{r}
set.seed(7035)
x_nonstandard <- rnorm(50,12,3); x_nonstandard
describe(x_nonstandard)
hist(x_nonstandard, main="Normal dist. mean=12, std=3")

```



Normality Test

```{r}
dt <- read.csv("C:\\Users\\user\\OneDrive\\Desktop\\Sem4 slide\\data science programming 2\\dt_clean.csv", header=TRUE)
dt
```

R code: Histogram and Q-Q PLot
Plot of histogram and fitted probability density function (pdf)

```{r}
dt$weight <- as.numeric(dt$weight)
```

```{r}
## Observed data
par(mfrow=c(2,1))
par(mar = c(3, 3, 2, 2))
hist(dt$weight, col="blue", border="black",
     prob = TRUE,xlab = "weight",
     main = "Weight: Histogram and PDF")

```

```{r}
plot(dt$weight)
lines(density(dt$weight),lwd = 2,col = "red")
qqnorm(dt$weight, main='Normal')
qqline(dt$weight, col="red")
```

R code: Shapiro-Wilk and Kolmogorov-Smirnov Tests

```{r}
## Shapiro-Wilk
shapiro.test(dt$weight)

```

```{r}
## Kolmogorov-Smirnov
ks.test(dt$weight, 'pnorm')


```






