---
title: "CHAPTER 9: VISUALISATION IN R"
output: html_notebook
---


```{r}
dt <- read.csv("C:\\Users\\user\\OneDrive\\Desktop\\Sem4 slide\\data science programming 2\\dt_clean.csv", header=TRUE)
dt
```


1. Bar Plot
• Bar plots are very useful to represent the count of any
categorical variables.

• Other related graphs for categorical data in R are spine plots
or mosaic plots.


Example 9.1 Bar Plot
```{r}
## Barplot
barplot(dt$weight, ylim=c(0,100),names=dt$ID_No,
ylab="weight(kg)", xlab="ID_No.",
main="Weight of Students", col=c(2,5,8))

barplot(dt$weight, xlim=c(0,100),
horiz=TRUE, names=dt$ID_No,
ylab="ID_No.", xlab="weight(kg)",
main="Weight of Students", col=c(2,5,8))

```


2. Dot Plot
• A Dot Plot is a graphical display of data using dots plotted on
a graph with an x-axis and y-axis.

• A dot plot is also known as a strip plot or dot chart.

• These types of charts are used to graphically depict certain
data trends or groupings.


Example 9.2 Dot Plot
```{r}
## Dot chart
dotchart(dt$weight, ylim=c(0,100),labels=dt$ID_No,
xlab="Weight(kg)", ylab="ID. No.",
main="Weight of Students", col='blue', pch=16, cex=1.2)
```


3. Pie Chart
• A pie chart, also known as a circle chart, is a circular diagram
that resembles a pie.

• A pie chart shows how a total amount is divided between
levels of a categorical variable as a circle divided into radial
slices.

Example 9.3 Pie Chart
```{r}
## Pie chart
# Calculate the percentage for each section

freq_section <- round(prop.table(table(dt$Section))*100, 2)

section_labels <- round(prop.table(table(dt$Section))*100, 2)

# Concatenate a ’%’ char after each value
section_labels <- paste(section_labels, "%", sep="")

pie(freq_section, labels=section_labels, col = rainbow(9),
    main="Percentage of Students by Section")

# Create a legend at the right
legend(1.5, 0.5, c("01G","02G","03G"), cex=0.8,
fill=rainbow(9))
```


4. Histogram
• Histogram shows how the data is distributed.

• It helps to understand the underlying statistical distribution of
the data.

• The shape of distribution of data can either be symmetric or
skewed.

• Also can detect the existence of outliers.

Example 9.4a Histogram
```{r}
## Histogram
hist(dt$weight, xlab="Weight(kg)", ylab="ID. No.",
main="Students weight", col="cyan")

```

Example 9.4b Histogram and Density
```{r}
x <- dt$weight
hist(x, freq = FALSE, main = "Histogram and density")

# Calculate density
dx <- density(dt$weight)

# Add density
lines(dx, lwd = 3, col = "red") ##ecdf
```


5. Box plots
• Box plots are designed to represent some characteristics of the
data like dispersion or skewness and also to identify outliers.

• Box plot is also known as box and whiskers plot.

• It presents five number summary which include minimum
value, quartiles (first, second and third) and the maximum
value.

• The location of medium in the box will determine the
skewness of the data distribution. If it is located in the
middle, then the data is symmetric. If the medium is close to
the first quartile, the data is right skewed. Otherwise, the
data is left skewed.

• The box plot also shows the existence of outliers.

Example 9.5a Box Plots
```{r}
## Box plot
boxplot(dt$weight~dt$Section,xlab="Section",ylab="Weight(kg)",
main="Section No. vs. Weight", col=c("blue","cyan","pink"))

```

Example 9.5b Notch Box Plots
```{r}
boxplot(dt$weight~dt$Gender,ylab="Gender",xlab="Weight(kg)",
main="Weight(kg) by Gender", col=c("cyan","pink"),
horizontal=TRUE, notch=TRUE, varwidth = TRUE)

```


6. Scatter Plots
• Scatter plots or is also known as dispersion graphs. The
scatter plots are used to check the relation between two
variables or three variables for 3D scatter plots.
• Scatter matrix or pairwise comparison can be used to check
the relationship between many variables.

Example 9.6a Scatter Plot
```{r}
## Scatter plots
plot(dt$weight,dt$bmi,xlab="Weight(kg)",ylab="BMI",
main="BMI vs. Weight", col="blue",pch=16, cex=1.5)
abline(lm(dt$bmi~dt$weight), col = "red", lwd = 2)
```
Example 9.6b 3D Scatter Plot
```{r}
## 3D scatter plot
library(scatterplot3d)
trees
attach(trees)
scatterplot3d(Girth, Height, Volume,
main = "3D Scatterplot of trees dataset")

```

Example 9.6c Scatter Plot Matrix
```{r}
#Scatter plot Matrices
pairs(trees, main = "Scatter plot matrix for trees data")

```


9.2 Graphical User Interface Development
• GUI is the acronym for graphical user interface, the interface
that allows users to interact with electronic devices, such as
computers, laptops, smartphones and tablets, through graphical
elements (icons, buttons, and menus).

• GUI is a software application programming for human-computer
interaction, replacing text-based commands with user-friendly
actions. Its goal is to present the user with decision points that
are easy to find, understand and use. In other words, GUI lets
you control your device with a mouse, pen or even your finger.

• To count as “successful” interactivity, the dashboard interface
design has to be human-centric.

```{r}

```




Dashboard and GUI Application
• In information technology (IT), the term dashboard refers to
a graphical user interface (GUI) that organizes and presents
information in a format that is easy to read and interpret.

• In business computer information systems, a dashboard is a
type of graphical user interface which often provides at-a-glance
views of key performance indicators (KPIs) relevant to a particular objective or business process. In other usage, ”dashboard”
is another name for ”progress report” or ”report” and considered a form of data visualization.



R packages to develop Dashboard and GUI
In R, there are several packages available for developing GUI and
dashboard applications. Here are a few popular ones:

1. RShiny: RShiny is a widely used R package for building web
applications with interactive user interfaces. It allows you to
create web-based GUIs and dashboards using R code. RShiny
provides a reactive programming framework that facilitates easy
integration of R code and interactive components such as sliders, input fields and plots.

2. RStudio Dashboard: RStudio provides its own set of packages for building dashboards. The flexdashboard package allows
you to create flexible and interactive dashboards using R Markdown. It supports various layout options, including multiple
panels, tabs, and responsive design. You can embed interactive
visualizations and data-driven content in your dashboards.

3. shinydashboard: The shinydashboard package is an extension
of the Shiny package and provides a framework for creating
dashboards with a more structured layout. It offers predefined
dashboard components such as sidebars, boxes, and tabs, which
help in organizing the content of your dashboard. You can easily
add interactive elements and visualizations to create dynamic
dashboards.

4. RGL: If you are looking for 3D graphics capabilities in your
GUI, the rgl package is a good option. It provides functions for
creating interactive 3D plots, scenes, and animations. You can
incorporate these 3D visualizations into your GUI or dashboard
application.


