---
title: "R Notebook"
output: html_notebook
---
Name: TEAN JIN HE     MatricID: SD21063

Exercise 1

Question a
```{r}
a <- 2.3
(6*a+42)/3^(4.2-3.62)
```
Question b
```{r}
log(12.43)
log10(12.43) #log to base 10
log2(12.43) #log to base 2
log(x=12.43, base=2) # alternative log to base 2
sqrt(12.43)
exp(12.43)
```
Question c
```{r}
d <- 20
r <- d/2
area_circle <- pi*(r^2)
area_circle
```
Question d
```{r}
weight <- c(69,63,57,55,47,48,66,50,49,45)
weight_sort = sort(weight, decreasing = FALSE)
weight_sort[weight_sort<50]
```

Exercise 2

Question a
```{r}
v1 <- seq(5,11, length=7);v1
v2 <- c(3,-1,4.5,-37)
rep(v2,times=2)
v3 <- sqrt(4+(2/3));v3
v4 <- c(v1,v2,v3);v4
```
Question b
```{r}
v5 <- v4[c(5,8,12)];v5
```
Question c
```{r}
length(v4)
x <- v4[-c(1,12)]
x
```
Question d
```{r}
v6 = sort(v4, decreasing = TRUE);v6
```
Question e
```{r}
x1 <- append(x,rep(7,times=2),after=2)
x2 <- append(x1,rep(10,times=4),after=7)
x3 <- append(x2,rep(-37,times=1),after=16)
x3

```

Exercise 3

Question a
```{r}
##file.choose()

dt <- read.csv("C:/Users/user/OneDrive/Desktop/Sem4 slide/data science programming 2/weight_height_BSD2223_asal2.csv", header = TRUE)
dt
```
Question b
```{r}
str(dt)
```
Question c
```{r}
head(dt,10)
tail(dt,10)
```
Question d
```{r}
dt1 <- dt[,-1]
dt1
```

Question e
```{r}
class(dt1[,c(4)])
class(dt1[,c(5)])
```
```{r}

transform(dt1,weight = as.numeric(weight)) 
class(dt1[,c(4)])
```
```{r}
# change the character columns to numeric
dt1$weight = as.numeric(as.character(dt1$weight))
dt1$height = as.numeric(as.character(dt1$height))
print(sapply(dt1, class))
```
```{r}
colSums(is.na(dt1))
dt1[3, 4] <- 55
dt1[59, 4] <- 55
dt1[32, 5] <- 1.77
dt1
```
```{r}
library(dplyr)
# if else condition of dataframe column  in R 
dt2 <- mutate(dt1, height = ifelse(height >= 100, height/100, height))
dt2
```
Question f
```{r}
# calculate BMI value

dt3 <- dt2 %>% mutate(Section, bmi = weight / (height ^ 2)) %>% select(Section:weight, height, bmi)
dt3
```
Question g
```{r}
# multiple if else condition of dataframe column  in R 

dt_clean <- mutate(dt3, categories = ifelse(bmi < 18.5, "Underweight", 
                                ifelse(bmi < 25, "Normal",
                                       ifelse(bmi < 30, "Overweight",
                                              ifelse(bmi < 35, "Obese", "Extremely Obese")))))
dt_clean
``````

Exercise 4

Question a
```{r}
C_degree <- function(F) (5/9)*(F-32)
C_degree
```
Question b
```{r}

F <- c(47,80,25,20,105,132,30,274,33,214,58,77)

vC <- sapply(F, C_degree)
round(vC, 2)
```

Exercise 5

```{r}
A <- c(2,1,-3)
B <- c(1,4,0,3,2,8)
C <- c(1,4,0,3,2,8,-1,5,0)
MatrixA <- matrix(A, nrow = 3, ncol = 1); MatrixA
MatrixB <- matrix(B, nrow = 3, ncol = 2); MatrixB
MatrixC <- matrix(C, nrow = 3, ncol = 3); MatrixC
```
Question a
```{r}
# 5a (3x1) != (3x2)
MatrixA %*% MatrixB
```
Question b
```{r}
t(MatrixA) %*% MatrixB
```
Question c
```{r}
t(MatrixB) %*% (MatrixA %*% t(MatrixA))
```
Question d
```{r}
(MatrixA %*% t(MatrixA)) %*% t(MatrixB)
```
```{r}
# 5e
((MatrixB %*% t(MatrixB))+(MatrixA %*% t(MatrixA))-(10/3))**-1
```

Exercise 6

Question a
```{r}
v1 <- seq(5.5,0.5)
Array1 <- array(v1, dim = c(4,2,6)); Array1
```
Question b
```{r}
Array1[3,2,4]
```
Question c
```{r}
Array1[2,,1]
```

Exercise 7

Question b
```{r}
load(file="dt_clean.RData")
dt1_clean
```
Question c
```{r}
dt1_clean$height 
```
Question d
```{r}
## by default right value inclusive

lbs <- c("height<1.5", "1.5<height<1.6", "height>1.6")
height_group <- cut(dt1_clean$height, breaks = c(0,1.5,1.6,Inf),labels=lbs)
height_group

levels(height_group) ## number of group
```
```{r}
## combine weight data with group
cbind(dt1_clean$height,height_group)
```
Question e
```{r}
## frequency in each group
table(height_group)

#OR

## Using factor
f_height_group <- factor(height_group, levels=c("height<1.5", "1.5<height<1.6", "height>1.6"),ordered=TRUE)
table(f_height_group)
```
```{r}
#freq. table of proportions for height variable
prop.table(table(height_group))
```

Question f
```{r}
## 2-way freq. table (by Section)
table(dt1_clean$Section, height_group)

##2-way freq. table (by gender)
table(dt1_clean$Gender, height_group)
```