---
title: "R Notebook"
output: html_notebook
---

example 4,1 create a list
```{r}
v1 <- seq(0,20,2) ## generate even numbers
chr1 <- c("Now I like R", "R is awesome")
logic1 <- c(TRUE, FALSE, TRUE, FALSE)
list_1 <- list(v1, chr1, logic1)
list_1

v2 <- seq(0,19,2) ## generate even numbers
m2 <- matrix(v2, nrow=2) ## matrix
chr2 <- c("Now I know R and Python", "R and Python are BFF")
logic2 <- c(TRUE, FALSE, TRUE, FALSE)
list_2 <- list(m2, chr2, logic1)
list_2

```
Example 4.2 - Naming List
```{r}
v2 <- seq(0,19,2) ## generate even numbers
m2 <- matrix(v2, nrow=2) ## matrix
chr2 <- c("Now I know R and Python", "R and Python are BFF")
logic2 <- c(TRUE, FALSE, TRUE, FALSE)
list_2 <- list(m2, chr2, logic2)

names(list_2) <- c("Matrix", "Character", "Logical")
list_2

```
Example 4.3 - Accessing List
```{r}
v2 <- seq(0,19,2) ## generate even numbers
m2 <- matrix(v2, nrow=2) ## matrix
chr2 <- c("Now I like R", "R is awesome")
logic2 <- c(TRUE, FALSE, TRUE, FALSE)
list_2 <- list(m2, chr1, logic1)

names(list_2) <- c("Matrix", "Character", "Logical")
list_2
```
Example 4.4 - Merging List
```{r}
##Example 4.4 - Merging List
merge_list <- c(list_1, list_2)
merge_list

merge_list[[2]]
merge_list[[5]]
merge_list$Character

#test
merge_list[[3]]
merge_list[[4]]
```
Example 4.5 - Deleting and Adding List
```{r}
## Deleting and Adding Element in the List
## Deleting element from the list
merge_list[[1]] <- NULL
merge_list

## Adding new element in a list
merge_list[[1]] <- "Updated information"
merge_list
```
Example 4.6 - Convert List to Vector
```{r}
## Convert List to Vector
v1 <- seq(0,20,2) ## generate even numbers
chr1 <- c("Now I like R", "R is awesome")
logic1 <- c(TRUE, FALSE, TRUE, FALSE)
list_1 <- list(v1, chr1, logic1)
list_1

unlist(list_1[1])

```
Example 4.7 - Special List
```{r}
## Special List
letters ##small letters
letters[1:5]

LETTERS ##capital letters
LETTERS[6:15]

month.abb ##month abbreviation
month.abb[3:12]

month.name ##month fullname
month.name[4:8]
```
Example 4.8 - apply(.) Function

Syntax: apply(X, MARGIN, FUN) and input data in the form of
data frame, matrix or a multi-dimensional array.
```{r}
m <- matrix((1:20), nrow=4, byrow=TRUE)
m

apply(m, 1, min) ## 1 refers to row
apply(m, 1, max)
apply(m, 1, mean)

apply(m, 2, sum) ## 2 refers to column
apply(m, 2, cumsum) ## cumulative sum
apply(m, 2, range)

## custom made function
mn2 <- function(x) mean(x)/2

apply(m, 1, mn2)
apply(m, 1, mean)

```
Example 4.9 - lapply(.) Function

Syntax: lapply(X, FUN) (l=list) and input data in the form of
data frame, vector and a list. The function applies element-wise
for lists and vectors and column-wise for data frames.

```{r}
## Function on List
x <- list(a = 1:5, beta = exp(-2:2),
logic = c(TRUE,TRUE,FALSE,FALSE, TRUE, FALSE))
x

# compute for each list element
lapply(x,sum) #sum
lapply(x,mean) #mean
lapply(x,median) #median
lapply(x,quantile) #quartiles

## custom made code
add_two <- function(x) x+2
lapply(x,add_two) #sum and add 2
```
Example 4.10 - sapply(.) Function

Syntax: sapply(X, FUN) (s=simplified) and input data in the
form of data frame, vector and a list. The function applies
element-wise for lists and vectors and column-wise for data
frames.

```{r}
## Function on List
v1 <- c(rep(2,3), seq(10,20, length=5))
v1

sapply(v1, mean) ## not working, element-wise
sapply(v1, add_two)
x <- list(a = 1:5, beta = exp(-2:2),
logic = c(TRUE,TRUE,TRUE,FALSE))
x

sapply(x, mean)
sapply(x, add_two)
```
Example 4.11 - List within a List
```{r}
## Recursive List
list_3 <- list(x3 = c(2,4), y3 = 1:5)
list_3
list_4 <- list(x4 = rep(1,2), y4 = seq(2,4, length=4))
list_4

new_list <- list(list_3, list_4)
new_list

new_list[[2]] ## extract list 2
new_list[[1]]$y3 ## extract list 1, element y3
new_list[[2]]$x4 ## extract list 2, element x4
```
Example 4.13 - Access List within a List
```{r}
## extract list 1, element y3
new_list[[1]]$y3
new_list[[1]][2]

## extract list 2, element x4
new_list[[2]]$x4
new_list[[2]][1]

## extract list 1, element 1 and element 1
new_list[[1]]$x3[1]
new_list[[1]][[1]][1]
new_list[[1]][[1]][2]

```
Example 4.14 - Access List within a List
```{r}
## extract list 2, element 2 and element 3
new_list[[2]][[2]][3]
## extract list 2, element 1 and element 1
new_list[[2]]$x4[1]
new_list[[2]][[1]][1]
## extract list 2, element 2 and element 3
new_list[[2]][[2]][3]
```

