---
title: "R Notebook"
output: html_notebook
---
Example 5.1 - Create a Data Frame
```{r}
## Create Data frame
v1 <- seq(2,10,length=4)
chr1 <- c("R","Python","Excel","Tableau")
logic1 <- c(TRUE,FALSE,TRUE,FALSE)
df_1 <- data.frame(v1, chr1, logic1)
df_1
str(df_1)
class(df_1)

Id_No <- c("SD2001","SD2002","SD2003","PE2007","PE2008")
Faculty <- c("PSM","PSM","PSM","FIM","FIM") ## character
Entry_level <- c("Matric","STPM","STPM","Matric","STPM")
BMI <- c(22.5, 24.2, 27.6, 30.7, 18.5)
df_2 <- data.frame(Id_No,Faculty,Entry_level,BMI)
df_2
str(df_2)
class(df_2)
data.frame()

```
Example 5.2 - Convert List to Data Frame
```{r}
## Convert list to Data frame
Id_No <- c("SD2001","SD2002","SD2003")
Faculty <- c("PSM","PSM","PSM") ## character
Entry_level <- c("Matric","STPM","STPM")
BMI <- c(22.5, 24.2, 27.6)
list_3 <- list(item1 = Id_No,item2 = Faculty,
item3=Entry_level,item4=BMI)
as.data.frame(list_3)
list_3 <- list(Id_No = Id_No,Faculty = Faculty,
Entry_level=Entry_level,BMI=BMI)
as.data.frame(list_3)

```
Example 5.3 - Convert Vector and Matrix
```{r}
## Vector to data frame
v2 <- seq(0,19,2) ## generate even numbers
v2 # vector
as.data.frame(v2) # vector to data frame

## Matrix to data frame
m2 <- matrix(v2, nrow=2) ## matrix
m2 # matrix
as.data.frame(m2) # matrix to data frame

```
Example 5.4 - Naming Data Frame
```{r}
m3 <- matrix(seq(1,20, by=2), nrow=5)
m3
df_3 <- as.data.frame(m3)
df_3
rownames(df_3) <- c("row1","row2","row3","row4","row5")
colnames(df_3) <- c("col1", "col2")
df_3
attributes(df_3)
class(df_3)

```
```{r}
##testing
#a.Create data frame with the following  criteria
#-4 variables
#-5 observation for each variable
#numeric, character

com <- c("Asus","ROG","MSI","Legion","Predator")
price <- c(2000,4000,4500,5000,3000)
country <- c("HK","TW","MY","CH","MY")
status <- c("work","gaming","gaming","gaming","gaming")
df4 <- data.frame(com,price,country,status)
df4
str(df4)
class(df4)


```
Built-in Data in R - mtcars
```{r}
data() ## all built-in data in R
data("mtcars")
dt <- mtcars
head(dt); tail(dt)
str(dt); class(dt)
```
Example 5.5 - Access & Extract Data using [ ]
```{r}
dt <- mtcars
dt_mpg <- dt[,1]
dt_mpg
dt_mpg <- dt$mpg
dt_mpg
# Extract row 1 to 15 and column 1,3 & 6
dt1 <- dt[1:15, c(1,3,6)] ; dt1
## Extract mpg > 25
dt[dt$mpg > 25, ]
## Extract 15 <mpg< 20
dt[dt$mpg > 15 & dt$mpg < 20, ]
```
Example 5.6 - Extract Data using subset(.) Function
```{r}
## subset function
subset(dt, mpg > 15 & cyl == 4, c("mpg", "cyl", "wt"))
subset(dt, (mpg >15 & mpg < 20) & cyl == 8,
c("mpg", "cyl", "wt"))
```
Example 5.7 - Merging Data Frame
```{r}
## Create the data frame.
birth_date1 <- data.frame(
id = c (1:5),
name = c("Ayra","Sara","Aisy","Aishah","Yusuf"),
height_m = c(1.2,1.1,1.3,1.3,1.1),
birth_date = as.Date(c("2013-12-28","2015-05-15",
"2015-05-02","2015-05-25","2016-12-13")),
stringsAsFactors = FALSE
)
birth_date1
birth_date2 <- data.frame(id = 6:9,
name = c("Warda","Lutfi","Mika","Alma"),
height_m = c(1,0.9,0.8, 0.2),
birth_date = as.Date(c("2017-08-31", "2019-09-30",
"2020-02-19", "2023-01-04")),
stringsAsFactors = FALSE
)
birth_date2
## Merge 2 data frames
## rbind(.) - row bind; cbind(.) - column bind
rbind(birth_date1, birth_date2)
## Add new column
all_birth <- rbind(birth_date1, birth_date2)
all_birth$gender <- c("F","F","M","F","M","F","M","M","F")
all_birth
```
Example 5.8 - apply(.) Function
```{r}
## apply(.) function on data frame
dt <- mtcars
apply(dt, 2, mean)
mean(dt$mpg)
apply(dt[,c(1,3,6)], 2, mean)

```
Example 5.9 - lapply(.) Function
```{r}
## lapply(.) function on data frame
dt3 <- airquality[,c(1:4)]
is.na(dt3)
sum(is.na(dt3))
which(is.na(dt3))
lapply(dt3, mean)
lapply(dt3, mean, na.rm=TRUE)

```
Example 5.10 - sapply(.) Function
```{r}
## sapply(.) function on data frame
dt3 <- airquality[,c(1:4)]
is.na(dt3)
sum(is.na(dt3))
which(is.na(dt3))
sapply(dt3, mean)
sapply(dt3, mean, na.rm=TRUE)
dt3[is.na(dt3)] <- colMeans(dt3,na.rm=TRUE)
dt3
dt3 <- airquality[,c(1:4)]
dt3[is.na(dt3)] <- 9999 # replace NA with 0
dt3
```
Example 5.11 Tibble versus Data Frame
```{r}
## Packages: tidyverse or tibble
## or install.packages("tibble")
install.packages("tidyverse")
library(tidyverse)
dt4 <- iris
head(dt4); tail(dt4)
str(dt4)
class(dt4)
dt4_tbl <- as_tibble(dt4)
dt4_tbl
dt5_tbl <- tibble(x = 1:10,y = 2, z = x ^ 2 - y/2)
dt5_tbl
```


